Pillar 1 - Sequential Pattern Mining

### 2019-02-12 Pillar 1 - Sequential Pattern Mining.Rmd
separate rhs and lhs in clean_AR

    # Clean variable `rules`
    mutate(
      rules = str_remove_all(rules, pattern = "[{]"),
      rules = str_remove_all(rules, pattern = "[}]")
      ) %>%
    separate(
      rules,
      into = c("lhs", "rhs"), 
      sep  = " => "
      )

@@

    # rhs.support
    mutate(
      rhs_course = str_remove_all(rhs, pattern = "_fail|_pass|_low|_not taken")
      ) %>%
    left_join(
      select(data_support, item, !!enquo(type_rule)),
      by = c("rhs_course" = "item")
      ) %>%
    mutate(
      rhs.support = !!enquo(type_rule)
      )

@@

  # Confidence and Lift by hand
  separate(
    rhs,
    sep = "_",
    into = c("rhs_course", "rhs_outcome"),
    remove = FALSE
    ) %>%

  group_by(
    lhs,
    rhs_course
    ) %>%
  mutate(
    lhs.rhsTake.support = sum(support) # Proba of lhs and taking rhs course
    ) %>%
  ungroup %>%

@@

  # Confidence and Lift
  separate(
    rhs,
    sep = "_",
    into = c("rhs_course", "rhs_outcome"),
    remove = FALSE
    ) %>%
  mutate(
    rhs.taken = str_detect(rhs, "fail|pass")
    ) %>%
  group_by(
    lhs,
    rhs_course
    ) %>%
    mutate(
    lhs.rhsTake.support = sum(support * rhs.taken) # Proba of lhs and taking rhs course
    ) %>%
  ungroup %>%

@@

  # separate rhs and lhs
  separate(
    rhs,
    into = c("rhs_course", "rhs_grade"),
    sep = "_",
    remove = FALSE,
    convert = TRUE
    ) %>%
  separate(
    lhs,
    into = c("lhs_course", "lhs_grade"),
    sep = "_",
    remove = FALSE,
    convert = TRUE
    ) %>%

  # rhs.support
  left_join(
    d_support$G,
    by = c("rhs_course" = "item", "rhs_grade" = "grade_ceil")
    ) %>%
  mutate(
    rhs.support = support.grade
    ) %>%

@@

  # reduce number of rules
  filter(
    lhs_course != rhs_course, # exclude rules with same course on lhs and rhs        
    rhs_grade == 6            
    ) %>%

@@

###TODO: fine tune parameters of cspade and ruleInduction, not taken --> fail

@@

clean_SR<- function(rules){

  rules %>%
    as("data.frame") %>%
    mutate(
      rule = str_remove_all(rule, pattern = "[<]"),
      rule = str_remove_all(rule, pattern = "[{]"),
      rule = str_remove_all(rule, pattern = "[}]"),
      rule = str_remove_all(rule, pattern = "[>]")
      ) %>%
    separate(rule, into = c("lhs", "rhs"), sep=" = ")  %>%
    select(
      lhs, rhs,
      support,
      confidence,
      lift
    ) %>%
    separate(lhs, into=c("LHS Course ID", "LHS Quality"), sep="_", remove= F) %>%
    separate(rhs, into=c("RHS Course ID", "RHS Quality"), sep="_", remove= F) %>%
    # Compute statistic
    group_by(lhs,`RHS Course ID`) %>%
    mutate(
      lhs.rhsCourse.support = sum(support),
      confidence = case_when(is.na(`RHS Quality`) ~ confidence,
                             T ~ (support / lhs.rhsCourse.support)),
      count = support * n_students,
      lhs.rhsCourse.count = lhs.rhsCourse.support * n_students,
      lift = 1 #dummy variable
      # TODO: compute rhs.support
      # lift = confidence.corr / rhs.support
     ) %>%
    ungroup() %>%
    mutate_at(
      c("support",
        "confidence", "lift"),
      funs(round(., 5))
    ) %>%
    filter(!`LHS Course ID`==`RHS Course ID`) %>%
    select(-`LHS Course ID`, -`LHS Quality`, -`RHS Course ID`, -`RHS Quality`)

}
```
We apply cspade followed by rule induction
```{r my_SR function}
# Run the cspade algorithm with desired parameters. Then run the ruleInduction algorithm with desired parameters.
my_SR<- function(data){
 data %>%
  cspade(
    parameter = list(
      #include as many item sequences as possible
      support = 0,   # min support of a sequence
      maxsize = 1,   # max number of items of an element of a sequence
      maxlen  = 2,   # max number of element of a sequence
      mingap  = 1,   # min time difference between consecutive element of a sequence
      maxgap  = 1e4  # max time difference between consecutive element of a sequence
      #maxwin  = 1e4 # max time difference between any two elements of a sequence # WARNING: 'maxwin' disabled
      ),
    control = list(
      verbose = FALSE
      )
    ) %>%
    ruleInduction(
      #include as many rules as possible
      confidence = 0,
      control    = list(verbose = FALSE)
      ) %>%
    clean_SR
}
```

>>>>

    # separate rule
    mutate(
      rules = str_remove_all(rules, pattern = "[{]|[}]")
      ) %>%
    separate(
      rules,
      into = c("lhs", "rhs"), 
      sep  = " => "
      ) %>%

    # separate rhs and lhs
    separate(
      rhs,
      into    = c("rhs_course", "rhs_outcome"),
      sep     = "_",
      remove  = FALSE,
      convert = TRUE
      ) %>%
    separate(
      lhs,
      into    = c("lhs_course", "lhs_outcome"),
      sep     = "_",
      remove  = FALSE,
      convert = TRUE
      )

@@

    # rhs.support
    left_join(
      select(data_support, item, !!enquo(type_rule)),
      by = c("rhs_course" = "item")
      ) %>%
    mutate(
      rhs.support = !!enquo(type_rule)
      )

@@

  # Confidence and Lift by hand
  group_by(
    lhs,
    rhs_course
    ) %>%
  mutate(
    lhs.rhsTake.support = sum(support) # Proba of lhs and taking rhs course
    ) %>%
  ungroup %>%

@@

  # Confidence and Lift
  mutate(
    rhs.taken = str_detect(rhs, "fail|pass")
    ) %>%
  group_by(
    lhs,
    rhs_course
    ) %>%
    mutate(
    lhs.rhsTake.support = sum(support * rhs.taken) # Proba of lhs and taking rhs course
    ) %>%
  ungroup %>%

@@

  # rhs.support
  left_join(
    d_support$G,
    by = c("rhs_course" = "item", "rhs_outcome" = "grade_ceil")
    ) %>%
  mutate(
    rhs.support = support.grade
    ) %>%

@@

  # reduce number of rules
  filter(
    lhs_course != rhs_course, # exclude rules with same course on lhs and rhs        
    rhs_outcome == 6            
    ) %>%

@@

```{r my_SR function}
# Run the cspade algorithm with desired parameters. Then run the ruleInduction algorithm with desired parameters.
my_cspade <- function(data){

  data %>%

    cspade(

      parameter = list(
        #include as many item sequences as possible
        support = 0,   # min support of a sequence
        maxsize = 1,   # max number of items of an element of a sequence
        maxlen  = 2,   # max number of element of a sequence
        mingap  = 1,   # min time difference between consecutive element of a sequence
        maxgap  = 1e4  # max time difference between consecutive element of a sequence
        #maxwin  = 1e4 # max time difference between any two elements of a sequence # WARNING: 'maxwin' disabled
        ),

      control = list(
        verbose = FALSE
        )

```{r}
my_ruleInduction <- function(rules){

  rules %>%

    ruleinduction(
      confidence = 0,
      control    = list(verbose = FALSE)
      )

@@

clean_SR <- function(rules){

  rules %>%

    as("data.frame") %>%

    # Isolate RHS and LHS
    mutate(
      rule = str_remove_all(rule, pattern = "[<]"),
      rule = str_remove_all(rule, pattern = "[{]"),
      rule = str_remove_all(rule, pattern = "[}]"),
      rule = str_remove_all(rule, pattern = "[>]")
      ) %>%

    separate(
      rule,
      into = c("lhs", "rhs"),
      sep = " = "
      ) %>%
    separate(
      lhs,
      into = c("LHS Course ID", "LHS Quality"),
      sep = "_",
      remove= FALSE
      ) %>%
    separate(
      rhs, 
      into = c("RHS Course ID", "RHS Quality"),
      sep = "_", 
      remove = FALSE
    )
}

    # Compute statistic
    group_by(lhs,`RHS Course ID`) %>%
    mutate(
      lhs.rhsCourse.support = sum(support),
      confidence = case_when(is.na(`RHS Quality`) ~ confidence,
                             T ~ (support / lhs.rhsCourse.support)),
      count = support * n_students,
      lhs.rhsCourse.count = lhs.rhsCourse.support * n_students,
      lift = 1 #dummy variable
      # TODO: compute rhs.support
      # lift = confidence.corr / rhs.support
     ) %>%
    ungroup() %>%
    mutate_at(
      c("support",
        "confidence", "lift"),
      funs(round(., 5))
    ) %>%
    filter(!`LHS Course ID`==`RHS Course ID`) %>%
    select(-`LHS Course ID`, -`LHS Quality`, -`RHS Course ID`, -`RHS Quality`)
